---
title: "Experiment report"
author: "Dusselier Group"
format:
  html: 
    toc: true
    self-contained: true
params:
  df: NA
  bd: NA
  gc: NA
  ms: NA
  path: NA
  material: NA
  preparation: NA
  cc: NA
  fcc: NA
  scc: NA
  chem_values: NA
  molar_flow: NA
  conversion: NA
  mass_balance: NA
  boxplot: NA
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F)

user_lib <- normalizePath(file.path(Sys.getenv("USERPROFILE"), "R", "win-library", as.character(getRversion())), winslash = "/")
.libPaths(c(user_lib, .libPaths()))

pkgs <- c(
  "reactable",
  "tidyverse",
  "reactablefmtr",
  "plotly",
  "dygraphs",
  "magrittr"
)
sapply(pkgs, \(x) {
  if (!requireNamespace(x)) {
    install.packages(x, repos = "https://cloud.r-project.org")
  } else {
    library(x, character.only = T)
  }
})
```


```{r}
compounds <- tibble::tribble(
  ~name,
  ~color,
  "Hydrogen",
  "orange",
  "Carbon monoxide",
  "lightgreen",
  "Carbon dioxide",
  "red",
  "Methane",
  "blue",
  "Ethane",
  "purple",
  "Ethylene",
  "yellow",
  "Propane",
  "gray",
  "Propylene",
  "royalblue",
  "n-butane",
  "deeppink",
  "i-butane",
  "mediumseagreen",
  "Cis-2-butene",
  "darkviolet",
  "t-2-butene",
  "chartreuse",
  "Nitrogen",
  "dodgerblue",
  "Argon",
  "turquoise",
  "Water",
  "olive"
)

change_str <- function(string, from, to, lower = F) {
  purrr::map_vec(
    .x = string,
    .f = ~ .x %>%
      {
        ifelse(lower, tolower(.), stringr::str_to_title(.))
      } %>%
      stringr::str_replace_all(pattern = from, replacement = to)
  )
}

custom_plot <- function(
  data,
  x,
  y,
  points = F,
  lines = F,
  hline = F,
  area = F,
  boxp = F,
  facet = NULL,
  xlab,
  ylab,
  title = NULL,
  scale_y = NULL,
  custom_color = F,
  args = list(),
  ...
) {
  p <- ggplot2::ggplot(data, ggplot2::aes(x = {{ x }}, y = {{ y }}, ...))

  p <- p +
    list(
      if (points) {
        if ("points" %in% names(args)) {
          do.call(ggplot2::geom_point, args$points)
        } else {
          ggplot2::geom_point()
        }
      },
      if (lines) {
        if ("lines" %in% names(args)) {
          do.call(ggplot2::geom_line, args$lines)
        } else {
          ggplot2::geom_line()
        }
      },
      if (area) {
        if ("area" %in% names(args)) {
          do.call(ggplot2::geom_area, args$area)
        } else {
          ggplot2::geom_area()
        }
      },
      if (hline) {
        if ("hline" %in% names(args)) {
          do.call(ggplot2::geom_hline, args$hline)
        } else {
          ggplot2::geom_hline()
        }
      },
      if (boxp) {
        if ("boxp" %in% names(args)) {
          do.call(ggplot2::geom_boxplot, args$boxp)
        } else {
          ggplot2::geom_boxplot()
        }
      },
      if (!is.null(facet)) {
        if ("facet" %in% names(args)) {
          purrr::exec(
            .fn = ggplot2::facet_wrap,
            ~ .data[[facet]],
            !!!args$facet
          )
        } else {
          ggplot2::facet_wrap(~ .data[[facet]])
        }
      },
      if (custom_color) {
        cmp <- compounds %>%
          {
            setNames(.$color, .$name)
          }
        ggplot2::scale_fill_manual(
          values = cmp[names(cmp) %in% unique(data$Compound)]
        )
      },
      if (!is.null(scale_y)) do.call(ggplot2::scale_y_continuous, scale_y)
    )

  p <- p +
    ggplot2::labs(x = xlab, y = ylab, title = title) +
    ggplot2::theme_bw()

  return(p)
}
```

```{css}
.dygraphs.html-widget.html-fill-item.html-widget-static-bound {
  overflow: hidden;
}

.plotly.html-widget.html-widget-static-bound.js-plotly-plot {
  height: auto !important; 
}
```

```{r}
custom_reactable <- function(bd, ...) {
  reactable(
    bd,
    defaultColDef = colDef(
      align = "center",
      headerStyle = list(background = "#f7f7f8")
    ),
    defaultPageSize = 10,
    onClick = "select",
    bordered = T,
    highlight = T,
    resizable = T,
    showSortable = T,
    ...
  )
}
```

::: {.callout-note}
**Sample name**: `r params$path %>% as.character %>% strsplit('/') %>% unlist %>% {.[length(.)]}`
:::

## Experiment description

### Leak test

```{r}
test <- params$df %>%
  as_tibble() %>%
  rowwise() %>%
  mutate(sum = sum(fic_110, fic_120, fic_130, fic_140)) %>%
  ungroup() %>%
  filter(n > 5 & p_set != 0 & sum == 0) %>%
  select(event, n, p_set, sum, pt_310) %>%
  filter(event == max(event)) %>%
  slice(c(1, n())) %>%
  {
    c(event = unique(.$event), dif = .$pt_310[1] - .$pt_310[2])
  }

if (!is.na(test[1])) {
  str_glue("Event: {test[1]} \nPressure difference: {round(test[2], 2)}")
} else {
  cat("No Leak test found in data")
}
```

### Catalyst characterization

:::{.callout-note}
**Material description:** `{r} params$material`

**Material preparation:** `{r} params$preparation`
:::


::: {.panel-tabset}

#### Details

```{r}
tryCatch(
  expr = {
    params$cc %>%
      as_tibble() %>%
      custom_reactable()
  },
  error = \(e) cat("Not information provided")
)
```

#### Fresh

```{r}
tryCatch(
  expr = {
    params$fcc %>%
      as_tibble() %>%
      filter(Value != "") %>%
      custom_reactable()
  },
  error = \(e) cat('Not information provided')
)
```

#### Spent

```{r}
tryCatch(
  expr = {
    params$scc %>%
      as_tibble() %>%
      filter(Value != "") %>%
      custom_reactable()
  },
  error = \(e) cat('Not information provided')
)
```

:::


---


### Log summary

```{r}
params$bd %>%
  as_tibble() %>%
  slice_head(n = 1, by = event) %>%
  select(event, date_time, name, time_duration, comment) %>%
  mutate(
    date_time = as.POSIXct(date_time, origin = "1970-01-01", tz = "UTC"),
    comment = dplyr::na_if(comment, ".na.character")
  ) %>%
  custom_reactable(
    columns = list(
      event = colDef(name = "Event", width = 100),
      date_time = colDef(name = "Start time", width = 200),
      name = colDef(name = "Event name", minWidth = 200),
      time_duration = colDef(name = "Duration", width = 100),
      comment = colDef(name = "Observations", width = 150)
    )
  )
```

## Raw data

### Composition plot

::: {.panel-tabset}

#### Time
```{r}
#| column: body-outset

plot <- params$gc %>%
  as_tibble() %>%
  select(-injection) %>%
  pivot_longer(cols = 5:ncol(.), names_to = "Compound", values_to = "value") %>%
  mutate(
    Compound = str_replace_all(Compound, "_", " ") %>% str_to_title(),
    time = as.numeric(time)
  ) %>%
  ggplot(aes(x = time, y = value, fill = Compound)) +
  geom_area(alpha = 0.6, color = "black", linewidth = 0.2) +
  geom_hline(yintercept = 100, linetype = "dashed") +
  labs(x = "Reaction time (min)", y = "Composition") +
  facet_wrap(~event, scales = "free", ncol = 2) +
  theme_bw() +
  theme(
    axis.text = element_text(color = "black", size = 10),
    axis.title = element_text(size = 12),
    panel.spacing = unit(1, "lines"),
    panel.background = element_rect(colour = "black"),
    plot.margin = unit(c(0, 0, 2, 0), "cm")
  )

total_height <- 180 * length(unique(params$gc$event))
ggplotly(plot, width = 900, height = total_height)
```

#### Temperature

```{r}
#| column: body-outset

plot <- params$gc %>%
  as_tibble() %>%
  select(-injection) %>%
  pivot_longer(cols = 5:ncol(.), names_to = "Compound", values_to = "value") %>%
  mutate(
    Compound = str_replace_all(Compound, "_", " ") %>% str_to_title(),
    time = as.numeric(time)
  ) %>%
  ggplot(aes(x = tic_300_pv, y = value, fill = Compound)) +
  geom_area(alpha = 0.6, color = "black", linewidth = 0.2) +
  geom_hline(yintercept = 100, linetype = "dashed") +
  labs(x = "Temperature (°C)", y = "Composition") +
  facet_wrap(~event, scales = "free", ncol = 2) +
  theme_bw() +
  theme(
    axis.text = element_text(color = "black", size = 10),
    axis.title = element_text(size = 12),
    axis.title.x = element_text(margin = margin(t = 15)),
    panel.spacing = unit(1, "lines"),
    panel.background = element_rect(colour = "black"),
    plot.margin = unit(c(0, 0, 2, 0), "cm")
  )

total_height <- 180 * length(unique(params$gc$event))
ggplotly(plot, width = 900, height = total_height)
```

:::

### Mass detector

```{r}
#| column: body-outset

if (is.null(params$ms)) {
  cat("No MS file found in folder")
} else {
  data <- params$ms %>%
    as_tibble() %>%
    select(
      time_absolute_date_time,
      `2_amu_hydrogen`,
      `18_amu_water`,
      `44_amu_co2_c3h8`
    ) %>%
    mutate(
      time_absolute_date_time = as.POSIXct(
        time_absolute_date_time,
        origin = "1970-01-01",
        tz = "UTC"
      ),
      across(.cols = contains("amu"), .fns = ~ smooth.spline(.x, spar = 0)$y)
    ) %>%
    rename_with(
      .fn = ~ str_replace_all(.x, "_", " "),
      .cols = contains("_amu_")
    )

  dygraph(data, xlab = "Reaction Time (min)") %>%
    dyRangeSelector() %>%
    dyLegend(width = 800) %>%
    dyOptions(useDataTimezone = TRUE) %>%
    dyOptions(strokeWidth = 2)
}

```

## Chemometrics

### Summary 
```{r}
#| column: body-outset

chem_values <- data.frame(params$chem_values) %>%
  dplyr::mutate(time = as.numeric(time))

chem_values %>%
  dplyr::slice_head(n = 1, by = event) %>%
  dplyr::select(1:5) %>%
  custom_reactable(
    columns = list(
      event = colDef(name = "Event", width = 100),
      name = colDef(name = "Name", width = 300),
      tecnique = colDef(name = "Technique", minWidth = 100),
      is = colDef(name = "Internal Standard", width = 200),
      qis = colDef(name = "QIS", width = 100)
    )
  )

```

### Plots

::: {.panel-tabset}

#### Molar flow

```{r}
#| column: body-outset

event_names <- chem_values %>%
  dplyr::summarise(name = unique(name), .by = event) %>%
  dplyr::mutate(name = stringr::str_c("Event: ", event, " - ", name)) %>%
  {
    setNames(.$name, .$event)
  }

molar_plot <- params$molar_flow %>%
  data.frame() %>%
  dplyr::mutate(time = as.numeric(time)) %>%
  custom_plot(
    x = time,
    y = value,
    fill = Compound,
    area = T,
    xlab = "Time (min)",
    ylab = "Molar flow (mol/h)",
    facet = "event",
    custom_color = T,
    args = list(
      area = list(alpha = 0.6, color = "black", linewidth = 0.2),
      facet = list(
        scales = "fixed",
        ncol = 2,
        labeller = ggplot2::as_labeller(event_names)
      )
    )
  ) +
  ggplot2::theme(
    axis.text = ggplot2::element_text(color = "black", size = 10),
    axis.title = ggplot2::element_text(size = 12),
    panel.spacing = ggplot2::unit(0.5, "cm"),
    plot.margin = ggplot2::unit(c(0, 0, 2, 2), "cm"),
    panel.background = ggplot2::element_rect(colour = "black")
  )

total_height <- 180 * length(unique(chem_values$event))
plotly::ggplotly(
  molar_plot,
  height = total_height,
  dynamicTicks = T,
  tooltip = "fill"
)
```

#### Conversion

```{r}
#| column: body-outset

event_names <- chem_values %>%
  dplyr::filter(technique == "TOS") %>%
  dplyr::summarise(name = unique(name), .by = event) %>%
  dplyr::mutate(name = stringr::str_c("Event: ", event, " - ", name)) %>%
  {
    setNames(.$name, .$event)
  }

conversion_plot <- params$conversion %>%
  data.frame() %>%
  dplyr::mutate(time = as.numeric(time)) %>%
  custom_plot(
    x = time,
    y = value,
    fill = Compound,
    shape = Compound,
    lines = T,
    points = T,
    xlab = "Time (min)",
    ylab = "Conversion (%)",
    facet = "event",
    scale_y = list(n.breaks = 10),
    args = list(
      facet = list(ncol = 2, labeller = ggplot2::as_labeller(event_names))
    )
  ) +
  ggplot2::theme(
    axis.text.y = ggplot2::element_text(color = "black", size = 10),
    axis.title.y = ggplot2::element_text(size = 12),
    panel.spacing = ggplot2::unit(0.5, "cm"),
    plot.margin = ggplot2::unit(c(0, 0, 2, 2), "cm"),
    panel.background = ggplot2::element_rect(colour = "black")
  )

total_height <- 180 + 180 * length(event_names) / 1.5
plotly::ggplotly(conversion_plot, height = total_height, dynamicTicks = T)
```

#### Mass balance

```{r}
#| column: body-outset

mass_plot <- params$mass_balance %>%
  data.frame() %>%
  dplyr::mutate(time = as.numeric(time)) %>%
  custom_plot(
    x = time,
    y = value,
    fill = Compound,
    lines = T,
    points = T,
    xlab = "Time (min)",
    ylab = "Mass balance",
    facet = "event",
    scale_y = list(n.breaks = 10),
    args = list(
      facet = list(ncol = 2, labeller = ggplot2::as_labeller(event_names))
    )
  ) +
  ggplot2::theme(
    axis.text.y = ggplot2::element_text(color = "black", size = 10),
    axis.title.y = ggplot2::element_text(size = 12),
    panel.spacing = ggplot2::unit(0.5, "cm"),
    plot.margin = ggplot2::unit(c(0, 0, 2, 2), "cm"),
    panel.background = ggplot2::element_rect(colour = "black")
  )

plotly::ggplotly(mass_plot, height = total_height, dynamicTicks = T)
```

```{r}
#| column: body-outset
#| warning: false

params$mass_balance %>%
  data.frame() %>%
  dplyr::group_by(event, Compound) %>%
  dplyr::summarise(avg = round(mean(value), 2), sd = round(sd(value), 2)) %>%
  custom_reactable(
    columns = list(
      event = reactable::colDef(name = "Event"),
      Compound = reactable::colDef(name = "Compound"),
      avg = reactable::colDef(name = "Average"),
      sd = reactable::colDef(name = "Standard deviation")
    ),
    style = "border-radius: '3px'"
  )

```

#### Box plot

```{r}
#| column: body-outset

p <- params$boxplot %>%
  data.frame() %>%
  custom_plot(
    x = Compound,
    y = value,
    fill = Compound,
    boxp = T,
    xlab = "",
    ylab = "Molar flow (mol/h)",
    facet = "event",
    args = list(
      facet = list(
        scales = "free",
        ncol = 2,
        labeller = ggplot2::as_labeller(event_names)
      ),
      boxp = list(show.legend = F)
    ),
    custom_color = T
  ) +
  ggplot2::stat_summary(
    fun = mean,
    geom = "point",
    shape = 23,
    fill = "white",
    size = 2.5
  ) +
  ggplot2::stat_summary(fun = min, geom = "point", shape = 25, size = 2.5) +
  ggplot2::stat_summary(fun = max, geom = "point", shape = 19, size = 2.5) +
  ggplot2::theme(
    axis.text.y = ggplot2::element_text(color = "black", size = 10),
    axis.title.y = ggplot2::element_text(size = 12),
    panel.spacing = ggplot2::unit(0.5, "cm"),
    plot.margin = ggplot2::unit(c(0, 0, 2, 2), "cm"),
    panel.background = ggplot2::element_rect(colour = "black"),
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )

boxp <- plotly::ggplotly(p, height = total_height)

boxp$x$data <- lapply(boxp$x$data, function(x) {
  if (x$type == "box") {
    x$q1 <- min(x$y)
    x$q3 <- max(x$y)
    x$boxpoints <- FALSE
    x$whiskerwidth <- 0
  }
  return(x)
})

boxp
```


:::
